---
title: "Untitled"
author: "CONSTANT YAOKUMAH"
date: "`r Sys.Date()`"
output: pdf_document
---

```{r, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
```

```{r}
library(dplyr)
library(tidyr)
library(lubridate)
library(stringr)
library(ggplot2)
```

## Q1
```{r}
Survey <- read.csv('https://www.lock5stat.com/datasets3e/StudentSurvey.csv', na.strings=c('',' ')) 
Survey_filter <- Survey %>% drop_na(Sex,Year)
Response_count <- Survey_filter %>% count(Sex, Year)
Response_counts <- Response_count %>%
  mutate(Year = factor(Year, levels = c("First Year", "Sophomore", "Junior", "Senior")))
Response_counts
Response_table <- Response_counts %>%
  pivot_wider(names_from = Year, values_from = n) 

colnames(Response_table) <- c("Gender", "First Year", "Sophomore", "Junior", "Senior")
Response_table


```


## Q2a
```{r}
url <- "https://raw.githubusercontent.com/dereksonderegger/444/master/data-raw/FlagMaxTemp.csv"
data <- read.csv(url)
dataMax  <- data %>% filter(Year == 2005) 

dataPivot <- dataMax %>% pivot_longer( 
    X1:X31,                # which columns to apply this to
    names_to  = 'Day',   # What should I call the column of old column names
    values_to = 'Temperature')

dataClean  <-dataPivot %>% pull(Day) %>% str_replace(pattern='X', replacement='') 

dataCleanTemp <- dataPivot %>% mutate(Day=dataClean, Year=as.character(Year), Month =as.character(Month))%>%drop_na() 
dataCleanTemp1 <-dataCleanTemp %>% mutate(Date=lubridate::ymd(paste(Year,Month,Day)))



dataCleanTemp1 %>% ggplot(aes(x=Date, y=Temperature)) +geom_line()
```
## Q2b 
 Create a line graph that gives the monthly average maximum temperature
        for 2013 - 2015. Again the x-axis should be the date and the axis
        spans 3 years.

```{r}

dataPivot <- data %>% filter(between(Year,2013,2015)) %>%pivot_longer( 
    X1:X31,                # which columns to apply this to
    names_to  = 'Day',   # What should I call the column of old column names
    values_to = 'Temperature') %>% group_by(Year,Month) %>% drop_na() 

dataCleanSum <- dataPivot %>% summarise(Average=mean(Temperature), .groups='drop') %>% mutate(
  Date = lubridate::ym(paste(Year,Month))
)

dataCleanSum %>% ggplot(aes(x=Date, y=Average)) +geom_line()

```



##Q4 Using cbind,rbind and join

```{r}
A <-  tribble(
~Name, ~Car,
'Alice', 'Ford F150',
'Bob',   'Tesla Model III',
'Charlie', 'VW Bug')

B <- tribble(
 ~First.Name, ~Pet,
 'Bob',  'Cat',
 'Charlie', 'Dog',
'Alice', 'Rabbit')
AB_cbind <- cbind(A,B)
AB_cbind

AB_join <- inner_join(A,B, by = c("Name"= "First.Name"))
AB_join

```


##Q4b Adding Alice pet Guinea Pig
```{r}
Alice_pet <- tribble(
  ~First.Name, ~Pet,
  'Alice', 'Guinea Pig'
)
Alice_pet

B <- bind_rows(B,Alice_pet)
B
```


## Q4c  Rejoining then with cbind
```{r}
#AB_cbind <- cbind(A,B)

#AB_cbind
```
Error in data.frame due to differing number of rows: A =3rows, B =4rows, resulting in dimension error
  


##Q 4ci  Using join
```{r}
AB_join <- inner_join(A,B, by = c("Name" = "First.Name"))
AB_join
```


## Q5 Questions

```{r}
Customers <- tribble(
      ~PersonID, ~Name, ~Street, ~City, ~State,
      1, 'Derek Sonderegger',  '231 River Run', 'Flagstaff', 'AZ',
      2, 'Aubrey Sonderegger', '231 River Run', 'Flagstaff', 'AZ',
      3, 'Robert Buscaglia', '754 Forest Heights', 'Flagstaff', 'AZ',
      4, 'Roy St Laurent', '845 Elk View', 'Flagstaff', 'AZ')
    
    Retailers <- tribble(
      ~RetailID, ~Name, ~Street, ~City, ~State,
      1, 'Kickstand Kafe', '719 N Humphreys St', 'Flagstaff', 'AZ',
      2, 'MartAnnes', '112 E Route 66', 'Flagstaff', 'AZ',
      3, 'REI', '323 S Windsor Ln', 'Flagstaff', 'AZ' )
    
    Cards <- tribble(
      ~CardID, ~PersonID, ~Issue_DateTime, ~Exp_DateTime,
      '9876768717278723',  1,  '2019-9-20 0:00:00', '2022-9-20 0:00:00',
      '5628927579821287',  2,  '2019-9-20 0:00:00', '2022-9-20 0:00:00',
      '7295825498122734',  3,  '2019-9-28 0:00:00', '2022-9-28 0:00:00',
      '8723768965231926',  4,  '2019-9-30 0:00:00', '2022-9-30 0:00:00' ) 
    
    Transactions <- tribble(
      ~CardID, ~RetailID, ~DateTime, ~Amount,
      '9876768717278723', 1, '2019-10-1 8:31:23',    5.68,
      '7295825498122734', 2, '2019-10-1 12:45:45',  25.67,
      '9876768717278723', 1, '2019-10-2 8:26:31',    5.68,
      '9876768717278723', 1, '2019-10-2 8:30:09',    9.23,
      '5628927579821287', 3, '2019-10-5 18:58:57',  68.54,
      '7295825498122734', 2, '2019-10-5 12:39:26',  31.84,
      '8723768965231926', 2, '2019-10-10 19:02:20', 42.83) 
    
    Cards <- Cards %>% 
      mutate( Issue_DateTime = lubridate::ymd_hms(Issue_DateTime),
              Exp_DateTime   = lubridate::ymd_hms(Exp_DateTime) )
    Transactions <- Transactions %>% 
      mutate( DateTime = lubridate::ymd_hms(DateTime))


```



## Q5a

```{r}

customer_name <- 'Derek Sonderegger'

Derek_Statement <- Customers %>%
  filter(Name == customer_name) %>%
  select(PersonID) %>%
  left_join(Cards, by = "PersonID") %>%
  select(CardID) %>%
  left_join(Transactions, by = "CardID") %>%
  left_join(Retailers, by = "RetailID") %>% 
  select(CardID,RetailID,DateTime, Amount, Name)

Derek_Statement

```

##Q5b Aubre
```{r}

#Aubrey's information
aubrey_info <- Customers %>%
  filter(Name == 'Aubrey Sonderegger')

#Aubrey's CardID and PersonID
cardID <- Cards %>%
  filter(PersonID == aubrey_info$PersonID) %>%
  pull(CardID)

personID <- aubrey_info$PersonID


#Close the existing card
Cards %>%
  filter(CardID == cardID) %>%
  mutate(Exp_DateTime = ymd_hms('2019-10-15 16:28:21'))

#new CardID
new_cardID <- '1234567890123456'

#new row for Aubrey's new card
new_card <- tibble(
  CardID = new_cardID,
  PersonID = personID,
  Issue_DateTime = ymd_hms('2019-10-15 16:28:21'),
  Exp_DateTime = ymd_hms('2022-10-15 16:28:21')
)

Cards <- rbind(Cards, new_card)
Cards


```


## Q5c

```{r}
# Temporary variables for the new transaction
card <- new_cardID
retailid <- 2
datetime <- ymd_hms('2019-10-16 14:30:21')
amount <- 4.98

#  the new transaction
Valid_Cards <- Cards %>% 
  filter(CardID == card, Issue_DateTime <= datetime, datetime <= Exp_DateTime)
Valid_Cards

if (nrow(Valid_Cards) == 1) {
  new_transaction <- tibble(
    CardID = card,
    RetailID = retailid,
    DateTime = datetime,
    Amount = amount
  )

  Transactions <- rbind(Transactions, new_transaction)
  Transactions
} else {
  print('Card Denied')
}

```

## Q5d
```{r}
card <- '9876768717278723'
retailid <- 2
datetime <- ymd_hms('2019-10-17 00:00:00')
amount <- 4.98  # Example nefarious amount

Valid_Cards <- Cards %>%
  filter(CardID == card, Issue_DateTime <= datetime, datetime <= Exp_DateTime)

if (nrow(Valid_Cards) == 2) {
  new_transaction <- tribble(
    ~CardID, ~RetailID, ~DateTime, ~Amount,
    card, retailid, datetime, amount
  )
  Transactions <- bind_rows(Transactions, new_transaction)
  Transactions
} else {
  print('Card Denied')
}

```

##Q5e
```{r}
customer_name <- 'Aubrey Sonderegger'

Aubrey_Statement <- Customers %>%
  filter(Name == customer_name) %>%
  select(PersonID) %>%
  left_join(Cards, by = "PersonID") %>%
  select(CardID) %>%
  left_join(Transactions, by = "CardID") %>%
  left_join(Retailers, by = "RetailID")  %>% 
  select(CardID,RetailID,DateTime, Amount, Name)

Aubrey_Statement
```






