---
title: "STA 445"
author: "CONSTANT YAOKUMAH"
date: "`r Sys.Date()`"
output: pdf_document
---

```{r}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
```

```{r}
library(ggplot2)
library(dplyr)
library(datasets)
library(tidyr)
library(ggrepel)
library(latex2exp)
```

## 1a. 
The `rownames()` of the table gives the country names and you should
create a new column that contains the country names. *`rownames`
```{r}
data('infmort', package = 'faraway')

infmort_data <- infmort %>% 
  mutate(Country = row.names(infmort))
infmort_data <- drop_na(infmort_data)
head(infmort_data)
```


## 1b
Create scatter plots with the `log10()` transformation inside the `aes()`command.
```{r}
ggplot(infmort_data, aes(x = log10(mortality), y = log10(income), color = oil)) + 
  geom_point() + facet_wrap(.~region) 

```


## 1c
 Create the scatter plots using the `scale_x_log10()` and `scale_y_log10()`.
Set the major and minor breaks to be useful and aesthetically pleasing.
Comment on which version you find easier to read.

```{r}
p2 <- ggplot(infmort_data, aes(x =mortality , y = income)) +
  geom_point(aes(color = oil)) +
  facet_wrap(.~ region) +
  labs(x = "Mortality Rate", y = " Income") +
  scale_x_log10(breaks = c(10, 100, 1000, 10000),labels = c("10", "100", "1,000", "10,000")) +
  scale_y_log10(breaks = c(100, 1000, 10000),labels = c("100", "1,000", "10,000"))

p2

```

In general, using the scale_x_log10() and scale_y_log10() functions is preferred because it provides better control over the axis scales and breaks. 
The labels on the axes will be in the original units, and it's easier to interpret the data.


## 1d
Select 10-15 countries to label and do so using the geom_text_repel() function.

```{r}
selected_countries <- sample(infmort_data$Country, 15)  # Select 10-15 countries to label

p3 <- p2 +
  geom_text_repel(data = subset(infmort_data, Country %in% selected_countries),
                   aes(label = Country))

p3

```


## 2a
Create a regression model for $y=$ `Volume` as a function of $x=$ `Height`.
```{r}
data(trees)
model <- lm(Volume ~ Height, data = trees)
data <- trees %>% mutate(fit = fitted(model))
head(data)
```

## 2b
Using the `summary` command, get the y-intercept and slope of the
regression line.
```{r}
summary(model)

coef(model)
```

## 2c
Using `ggplot2`, create a scatter plot of Volume vs Height. 
```{r}
library(ggplot2)

p <- ggplot(data, aes(x = Height, y = Volume)) +
  geom_point() +
  labs(x = "Height", y = "Volume", title = "Volume vs Height")
p
```


## 2d
 Create a nice white filled rectangle to add text information to using by
adding the following annotation layer
```{r}
p1 <- p + annotate(
  "rect", xmin = 65, xmax = 75, ymin = 60, ymax = 74,
  fill = "white", color = "black"
)
p1
```

## 2e
Add some annotation text to write the equation of the line
```{r,warning=FALSE}

p2 <- p1 + annotate(
  "text", x = 70, y = 68,
  label = latex2exp::TeX('$\\hat{y}_{i} = -87.12 + 1.54 * x_{i}$'),
  
)
p2
```


## 2f
Add annotation to add $R^2 = 0.358$
```{r,warning=FALSE}
p3 <- p2 + annotate(
  "text", x = 70, y = 72,
  label = latex2exp('$R^{2} = 0.358$')
)
p3
```

## 2g
Add the regression line in red. The most convenient layer function to uses
is `geom_abline()`
```{r, warning=FALSE}
p4 <- p3 + geom_abline(intercept = coef(model)[1], slope = coef(model)[2], color = "red")
p4
```


