---
title: "STA 445"
author: "Constant Yaokumah"
date: "`r Sys.Date()`"
output: pdf_document
---

```{r, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
```


```{r}
library(stringr)
library(refinr)
library(dplyr)
library(lubridate)
library(mosaicData)

```

## Q1
```{r}
strings <- c("seed", "banana", "pens", "pineapple")
data.frame( string = strings ) %>%
  mutate( result = str_detect(string, 'a') )
```
Return TRUE if  "strings" contain 'a' in any word, otherwise FALSE


## Q1b
```{r}
strings <- c("able", "banana", "I am able", "pineapple")
data.frame( string = strings ) %>%
  mutate( result = str_detect(string, 'ab') )
```
Return TRUE if " strings" contain 'ab' in  that order in any word, otherwise FALSE

## Q1 c

```{r }
strings <- c("able", "banana", "pepper", "pineapple")
data.frame( string = strings ) %>%
  mutate( result = str_detect(string, '[ab]') )

```
Return TRUE if " strings" contain 'ab' ,'b' or 'a', order doesn't matter in any word, otherwise FALSE

## Q d
```{r}
strings <- c("able", "banana", "peppera", "pineapple")
data.frame( string = strings ) %>%
  mutate( result = str_detect(string, '^[ab]') )
```
Return TRUE if " strings" contain 'ab' ,'b' or 'a',must begin the word, order doesn't matter in any word, otherwise FALSE

## Q e
```{r }
strings <- c(" 457 Aaron ", "banana", "pepper"," 2345 Amen ")
data.frame( string = strings ) %>%
   mutate( result = str_detect(string, '\\d+\\s[aA]') )
```
Return TRUE if "strings" matches one or more digits, any white space and either lowercase letter 'a' or uppercase letter 'A', otherwise FALSE
## Q f

```{r}
strings <- c(" 457 Aaron ", "banana", "pepper"," 2345   Amen ","5678 a")
data.frame( string = strings ) %>%
   mutate( result = str_detect(string, '\\d+\\s*[aA]') )

```

Return TRUE if "strings" matches one or more digits,zero or more white space and either lowercase letter 'a' or uppercase letter 'A', otherwise FALSE

## Q g
```{r}
strings <- c("Aaron457  ", "banana", "pepper"," 2345   Amen ")
data.frame( string = strings ) %>%
  mutate( result = str_detect(string, '.*') )
```
Return TRUE if "strings" contain any character with zero or more repetitions of previous, otherwise FALSE

## Q h
```{r}
strings <- c("aybar", "cdbar", "5abbar","Kabbar")
data.frame( string = strings ) %>%
   mutate( result = str_detect(string, '^\\w{2}bar') )
```
Return TRUE if "strings" matches the start of a string, any alphanumeric character and exactly two word character expected with literal string 'bar', otherwise FALSE


## Q i
```{r}
strings <- c("foo.bar","xybar","Kabbar")
data.frame( string = strings ) %>%
  mutate( result = str_detect(string, '(foo\\.bar)|(^\\w{2}bar)') )

```
Return TRUE if "strings" matches 'foo.bar' or   matches the start of a string, any alphanumeric character and exactly two word character expected with literal string 'bar', otherwise FALSE

## Q2 
```{r}
 file.names <- c('S123.P2.C10_20120621_213422.jpg',
   'S10.P1.C1_20120622_050148.jpg',
   'S187.P2.C2_20120702_023501.jpg')
file.name <- str_replace_all(file.names, pattern = '_',replacement = '.')
file.name.split <- data.frame(str_split_fixed(file.name, pattern = '\\.', n = 6))
year_mon_day <- (file.name.split[,4])
hour_min_sec <- (file.name.split[,5])
Year <- str_sub(year_mon_day, start = 1, end = 4)
Month <- str_sub(year_mon_day, start = 5, end = 6)
Day <- str_sub(year_mon_day, start = 7, end = 8)
Hour <- str_sub(hour_min_sec, start = 1, end = 2)
Minute <- str_sub(hour_min_sec, start = 3, end = 4)
Second <-  str_sub(hour_min_sec, start = 5, end = 6)
Site <- file.name.split[,1]
Plot <- file.name.split[,2]
Camera <- file.name.split[,3]
file.name.final <- data.frame(Site,Plot,Camera,Year,Month,Day,Hour,Minute,Second)
file.name.final

```


## Q3
```{r}
Gettysburg <- ('Four score and seven years ago our fathers brought forth on this 
continent, a new nation, conceived in Liberty, and dedicated to the proposition 
that all men are created equal.

Now we are engaged in a great civil war, testing whether that nation, or any 
nation so conceived and so dedicated, can long endure. We are met on a great 
battle-field of that war. We have come to dedicate a portion of that field, as 
a final resting place for those who here gave their lives that that nation might 
live. It is altogether fitting and proper that we should do this.

But, in a larger sense, we can not dedicate -- we can not consecrate -- we can 
not hallow -- this ground. The brave men, living and dead, who struggled here, 
have consecrated it, far above our poor power to add or detract. The world will 
little note, nor long remember what we say here, but it can never forget what 
they did here. It is for us the living, rather, to be dedicated here to the 
unfinished work which they who fought here have thus far so nobly advanced. It 
is rather for us to be here dedicated to the great task remaining before us -- 
that from these honored dead we take increased devotion to that cause for which 
they gave the last full measure of devotion -- that we here highly resolve that 
these dead shall not have died in vain -- that this nation, under God, shall 
have a new birth of freedom -- and that government of the people, by the people, 
for the people, shall not perish from the earth.')

words <- str_replace_all(Gettysburg, pattern = '\\-\\-|\\-|\\.|\\,', replacement = '')
words <- str_split(words, "\\s+")[[1]]
length_word <- nchar(words)
mean_word_length <- mean(length_word)
paste("Mean_word_length:",mean_word_length)


```



## 12 Q1
```{r}
library(lubridate)

mdy("September 13, 2010.")
```

```{r}
mdy("Sept 13, 2010")  #b) Sept 13, 2010.
```

Return NA because the Sept is not the default short of September

```{r}
mdy("Sep 13, 2010.")    #c) Sep 13, 2010.
```

```{r}
mdy("S 13, 2010.")      #d) S 13, 2010. Comment on the month abbreviation needs.
```
Return NA because the S is not the default short of September")

```{r}
dmy("07-Dec-1941.") #e) 07-Dec-1941.
```

```{r}
mdy("1-5-1998")#f) 1-5-1998. Comment on why you might be wrong.
```
This might be wrong because the month could be the day and vise visa


```{r}
dmy("21-5-1998")#g) 21-5-1998. Comment on why you know you are correct.
```

```{r}
ymd_hm("2020-May-5 10:30 Am")
```

```{r}
ymd_hm("2020-May-5 10:30 Am", tz = "America/Los_Angeles")
```

```{r}
ymd_hm("2020-May-5 10:30 Am", tz = "America/Puerto_Rico")   
  
```


## Q2
```{r}

birth_date <- ymd("1998-09-07")

current_date <- Sys.Date()

birthday_64th <- birth_date + years(64)

age_years <- ceiling(as.numeric(difftime(current_date, birth_date, units = "days") / 365))-1

next_birthday <- birth_date + years(age_years + 1)

days_until_next_birthday <- as.numeric(difftime(next_birthday, current_date, units = "days"))

months_until_next_birthday <- as.integer(days_until_next_birthday / 30)

days_remaining_until_next_birthday <- days_until_next_birthday %% 30



paste("birth_date    ",birth_date)
paste("current_date    ",current_date)
paste("birthday_64th    ",birthday_64th)
paste("age_years   ",age_years)
paste("next_birthday   ",next_birthday)
paste("days_until_next_birthday    ",days_until_next_birthday)
paste("months_until_next_birthday   ",months_until_next_birthday)
paste("days_remaining_until_next_birthday ",days_remaining_until_next_birthday)

```


## Q 3
```{r}

Arizona_time <- with_tz(ymd_hm("2015-05-08 15:00 ", tz = "America/Phoenix"), "America/Phoenix")

Auckland_time <- with_tz(Arizona_time, "Pacific/Auckland")
Arizona_time
Auckland_time

```

## Q5

```{r}
library(mosaicData)
library(ggplot2)
library(dplyr)

data(Births78)
Birth_78 <- subset(Births78, select = c(date, births))
ggplot(Birth_78, aes(x = date, y = births)) + geom_point() + labs(x = "Date" , y = "Number of Births", title = " Number od Births and the Dates ")

Birth_78 <- Birth_78 %>% mutate(dow = wday(date, label = TRUE))
ggplot(Birth_78, aes(x = date, y = births, color = dow )) + geom_point() + labs(x = "Date" , y = "Number of Births", title = " Number of Births in days of the weeks")

```








