---
title: "STA 445"
author: "Constant Yaokumah"
subtitle: "6260662"
date: "`r Sys.Date()`"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Q8. 1 create vectors,label and sum
```{r, }
vec_a <- c(2,4,6)
vec_b <- c(8,10,12)
vec_c <- vec_a + vec_b
vec_c
```
## Q8. 2 Add vectors
```{r}
vec_d <- c(14,20)
vec_d + vec_a
```
The vector d = (14,20), however, R recreate the vector d as (14,20,14) because the length of d did not correspond to vector a 
Warning: longer object length is not a multiple of shorter object length[1] 16 24 20


## Q8. 3 Add 5
```{r }
vec_a + 5   # R refers to the 5 as a scalar added to each vector in vec_a
```
When 5 was added, R create the 5 as a scalar adding to each element in the vector a, hence no warning 


## Q8. 4  Generating vectors
```{r }
seq(1,5)  #a
1:5    #b
```

## Q8. 5 Generating vectors of even numbers
```{r,}
seq(2,20,2)  #a
2*(1:10)   #b
```

## Q8. 6 vector of 21 elemnet from 0 to 1 
```{r}
x <- seq(from =0, to = 1, length.out = 21)
x
```

## Q8. 7 Using rep 
```{r}
rep(c(2,4,8),3)
```

## Q8. 8  Using rep for each
```{r}
rep(c(2,4,8), each= 4)
```


## Q8. 10 Matrix 
```{r}
M = matrix(seq(2,30,2), nrow = 3, ncol = 5, byrow = TRUE) #a
M
## matrix using seq and cbind
rbind(seq(2,10,2), seq(12,20,2),seq(22,30,2)) #a i

M[2,] #b

M[3,2]
```

## Q8. 12 Dataframe and deleting NA
```{r }
df <- data.frame(name= c('Alice','Bob','Charlie','Daniel'),
                     Grade = c(6,8,NA,9))
    
df[ -which(  is.na(df$Grade) ), ] 
df[  which( !is.na(df$Grade) ), ]

```
 Both copes perform the same work, however, the difference lies on how the copes are written , the first (is.na) select any (NA) and (-which) removes or delete that row while the second is (!is.na) removes the row which any (NA) 
 
## Q8. 14 List named of element
```{r}
x = c(4,5,6,7,8,9,10)
y = c(34,35,41,40,45,47,51)
slope = 2.82
p.value = 0.000131

my.test <- list(x = x,y = y,slope = slope,p.value = p.value)
my.test #a
my.test[2] #b

my.test$p.value   # my.test['p.value'] c
```

## Q 9 1 Example_5
```{r}
library(readxl)
Example_5 <- read_excel('Example_5.xls', sheet ='RawData',range = "A5:C36")
head(Example_5)  
str(Example_5)
```


## Q 9 2 Example_3
```{r }
Example_3 <- read_excel('Example_3.xls',sheet = 'data', range = "A1:L34", na = c('NA', '-9999'))
tail(Example_3) # 
```





